---
title: "Forecast Methods"
author: "Pawan Kumar"
date: "02/06/2021"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Average method

Forecast at time step "h" is average of the previous historic data

$y_{t+h|t}$ = ( $y_1$ + $y_2$ +........$y_t$ ) / t

```{r}
# mean(y,h) -> y is the time series and h is the forecast horizon

library(fpp2)
beer_data <- window(ausbeer, start = 1992, end = c(2007,3))
beer_data[1:20]
frequency(beer_data) 
```

| Frequency | Meaning       |
|-----------|---------------|
| 12        | Monthly       |
| 4         | Quarterly     |
| 2         | Semi annually |
| 1         | Annual        |

: Frequency table

```{r cars}
summary(beer_data)

```

## Plot

```{r pressure, echo=FALSE}

autoplot(beer_data) +
autolayer(meanf(beer_data, h=11),
series="Mean", PI=FALSE) +
autolayer(naive(beer_data, h=11),
series="Naïve", PI=FALSE) +
autolayer(snaive(beer_data, h=11),
series="Seasonal naïve", PI=FALSE) +
ggtitle("Forecasts for quarterly beer production") +
xlab("Year") + ylab("Megalitres") +
guides(colour=guide_legend(title="Forecast"))
```

## Box-Cox Transformation

Usage of BoxCox.lambda() function to choose adequate lambda value:

```{r}

lambda_value <- BoxCox.lambda(elec)
lambda_value

autoplot(BoxCox(elec,lambda_value))
```

```{r}
transformed_series <- BoxCox(elec,lambda_value)
# Plotting
autoplot(transformed_series)+
autolayer(meanf(transformed_series,h=40), series = "Mean", PI = FALSE) +
autolayer(naive(transformed_series, h = 40), series = "Naive", PI = FALSE) +
autolayer(snaive(transformed_series,  h = 40), series = "Seasonal", PI = FALSE)+
ggtitle("Forecasted electricity demand") +
xlab("Time") +
ylab("Power-volt") +
guides(color = guide_legend(title = "Forecast"))
  
  

```

Note: Mean and naive forecast make no sense, to an extent the seasonal
forecast aligns, still it lacks to capture the trend. As the transformed
series has trend and seasonality.

Importantly, after chosing a transformation, we need to revert it back
to reflect upon the original data and generate and thus gets its
forecast value.

```{r}

seasonal_forecast <- snaive(transformed_series, h = 11)

View(seasonal_forecast)
length(seasonal_forecast[["lower"]])

# lower values
reverse_transformation <- InvBoxCox(seasonal_forecast[["lower"]][1:22], lambda_value)
```

## Bias Adjusted Means

```{r}
View(eggs)

fc1 <- rwf(eggs, h = 50, drift = TRUE)
fc2 <- rwf(eggs, h = 50, drift = TRUE, biasadj = TRUE)

autoplot(eggs) +
autolayer(fc1, series = "Simple back transformation") +
autolayer(fc2, series = "Bias adjusted") +
ggtitle("Bias adjusted comparison chart") +
guides(color = guide_legend(title = "Forecast"))



```
